Cognate beginnings to bilingual lexical acquisition

### 2023-07-06 manuscript/manuscript.tex
Fix appendix

source of this study. GGC was supported by the This research was
supported by a FPI 2020 research contract (PRE2019-088165) in a project
funded by the Spanish Ministry of Science, Innovation, and Universities
(PGC2018-101831-B-I00). DAV was supported by the European Union's
Horizon 2023 research and innovation program under Marie
Skłodowska--Curie Grant (765556) and a postdoctoral fellowship from the
Foundation for Science and Technology of Portugal (UIDB/00214/2020). NSG
was supported by Academia 2019 award from the Catalan Institution for
Research and Advanced Studies (ICREA). We are grateful to Chiara
Santolin, Ege E. Özer, Alicia Franco-Martínez, Cristina Rodríguez-Prada,
and the rest of the Speech Acquisition and Perception research group for
their helpful feedback. We thank Xavier Mayoral, Silvia Blanch, and
Cristina Cuadrado for their technical support, and Cristina Dominguez
and Katia Pistrin for their efforts in recruiting infants. We also thank
all families and infants who participated in the experiments.}}

@@

08005\\\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}\And

@@

08005\\\href{mailto:avila.varela.daniela@gmail.com}{avila.varela.daniela@gmail.com}\And

@@

28049\\\href{mailto:jignaciocastillejo@gmail.com}{jignaciocastillejo@gmail.com}\And

@@

08005\\\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, sharp corners, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden]}{\end{tcolorbox}}\fi

@@

Consider the example of the Catalan-Spanish cognate translation
equivalent /\textipa{"gat}-\textipa{"ga.to}/ {[}cat{]}, that are
phonologically very similar. When the child is exposed to
/\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in
parallel. Therefore, this exposure will count as a learning instance for
both co-activated forms. The case of the non-cognate translation
equivalent /\textipa{"gos-"pe.ro}/ would be different. Given the low
phonological similarity between both word-forms, an exposure to
\textipa{"gos} will result in a weak activation of /\textipa{"pe.ro}/
leading to such exposure counting as a learning instance for
/\textipa{"gos}/ (which the child was exposed to), but not for
/\textipa{"pe.ro}/. While /\textipa{"gat}-\textipa{"ga.to}/ will benefit
from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will not. If
the child receives linguistic input from one of the languages more often
than from the other, this effect might affect each form of the cognate
translation equivalent differently. For instance, if the child receives
a larger amount of Catalan input than Spanish input, they will encounter
the Catalan form /\textipa{"gat}/ more frequently than the Spanish form
/\textipa{"ga.to}/. Through parallel activation, /\textipa{"gat}/ will
activate /\textipa{"ga.to}/ more often than vice versa. Ultimately,
/\textipa{"ga.to}/ will benefit more strongly from its cognate status
than /\textipa{"gat}/, as it receives additional learning instances from
its translation equivalent more often than /\textipa{"gat}/.

@@

\centering
\begin{tabular}{lccccl}
\toprule
Semantic category & List A & List B & List C & List D & Examples\\
\midrule
Household items & 31 & 26 & 30 & 25 & clock, video\\
Food and drink & 29 & 26 & 23 & 27 & sausage, yogurt\\
Animals & 26 & 23 & 19 & 25 & panther, tiger\\
Outside & 14 & 13 & 13 & 15 & farm , stone\\
Body parts & 14 & 12 & 11 & 11 & face  , finger\\
\addlinespace
Toys & 11 & 11 & 12 & 13 & piano , racket\\
Clothes & 12 & 12 & 10 & 10 & zipper, sandal\\
Vehicles & 9 & 10 & 11 & 10 & helicopter, tractor\\
Colours & 6 & 6 & 6 & 6 & red  , green\\
People & 7 & 4 & 6 & 6 & police    , babysitter\\
\addlinespace
Furniture and rooms & 4 & 4 & 4 & 4 & corridor, terrace\\
Time & 2 & 2 & 2 & 2 & day  , night\\
Adventures & 1 & 1 & 1 & 1 & witch\\
Parts of things & 1 & 1 & 1 & 1 & wheel\\
\midrule
N & 167 & 151 & 149 & 156 & \\
\bottomrule
\end{tabular}
\end{table}

@@

We collected 138,078 responses to 1,590 items. We restricted the
analyses to responses to nouns (628 items corresponding to other
grammatical classes were excluded, see
\protect\hyperlink{ref-fourtassi2020growth}{Fourtassi et al., 2020} for
a similar approach). We then excluded items with missing lexical
frequency scores (\emph{n} = 269), items that included more than one
lemma (e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} = 48), multi-word
items or phrases (e.g., \emph{barrita de cereales} {[}cereal bar{]},
\emph{n} = 9). Finally, we removed items without a translation in the
other language (\emph{n} = 32). This resulted in a final list of 604
items, corresponding to 302 Catalan words and their 302 Spanish
translations (302 translation equivalents). After collecting
participants' responses, the final dataset consisted of 138,078
observations, each corresponding to a single response of one participant
to one item. Each translation equivalent received a median of 234
responses (\emph{Min} = 106, \emph{Max} = 872) from participants, both
languages pooled together.

@@

We modelled the probability of participants answering each response
category (\emph{No} \textless{} \emph{Understands} \textless{}
\emph{Understands and Says}) using a Bayesian, multilevel, ordinal
regression model. This model allowed us to estimate both item and
participant word-acquisition trajectories, while estimating the effect
of our variables of interest: \emph{Age} (number of months elapsed
between participants' birth date and questionnaire completion),
\emph{Length} (number of phonemes in the X-SAMPA phonological
transcription of the word-form), \emph{Exposure} (a language
exposure-weighted lexical frequency), and \emph{Cognateness} (defined as
the phonological similarity between translation equivalents). We added
these variables as main effects, together with the two-way and three-way
interactions between \emph{Age}, \emph{Exposure}, and
\emph{Cognateness}. Participant-level and item-level random intercepts
and slopes were included where appropriate, according to the structure
of the data (\protect\hyperlink{ref-barr2013random}{Barr et al., 2013}).
We specified a weakly informative prior around the parameters of the
model. Equation~\ref{eq-linear} shows a detailed description of the
model.

\begin{equation}\protect\hypertarget{eq-linear}{}{
\begin{aligned}
&\textbf{Response}~(k)~\textbf{to word}~i~ \textbf{by participant}~j \\
\text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}) \\ 
&\text{where}~k \in \{\text{No} \rightarrow \text{Understands}, \text{Understands} \rightarrow \text{Understands and Says}\}\\
&\textbf{Distribution parameters} \\
\theta_{k_{ij}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + \\
&(\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + 
(\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{ij} + (\beta_{5} + u_{5_{i}} + w_{3_{j}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
& (\beta_{6} + u_{6_{i}}) \cdot (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
& (\beta_{7} + u_{7_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
& (\beta_{8} + u_{8_{i}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij}) \\
&\text{where:}\\
& u_{1-8_{i}}~\text{: participant-level adjustments} \\
& w_{1-3_{j}}~\text{: TE-level adjustments} \\
&\textbf{Prior} \\
\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.5); ~\beta_{1-5} \sim \mathcal{N}(0, 1) \\
\sigma_{u_{0-8}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2) \\
&\text{where }~\rho_{u_{0-8}, w_{0-3}}~\text{are the correlations between group-level adjustments}\\
\end{aligned}
}\label{eq-linear}\end{equation}

@@

We developed the \emph{Exposure} predictor to account for the fact that
bilinguals' exposure to a given word-form is not only a function of the
word-forms lexical frequency, but also of the quantitative input they
receive from the language such word-form belongs to, we expressed
lexical frequencies as the product between both variables. First, we
extracted the child-directed lexical frequency of each word from the
CHILDES database
(\protect\hyperlink{ref-macwhinney2000childes}{MacWhinney, 2000};
\protect\hyperlink{ref-sanchez2019childesdb}{Sanchez et al., 2019}).
Using the corresponding lexical frequencies directly from Catalan and
Spanish was not possible due to the low number of Catalan participants
and tokens available in their corresponding CHILDES corpora, so they
were extracted from the English corpora instead. We then mapped the
lexical frequencies of the English words to their Catalan and Spanish
translations (see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
et al., 2020} for a similar approach), and transformed them to Zipf
scores (\protect\hyperlink{ref-vanheuven2014subtlexuk}{Van Heuven et
al., 2014}; \protect\hyperlink{ref-zipf1949human}{Zipf, 1949}). We then
multiplied the resulting lexical frequencies by the reported degree of
exposure of the child to Catalan or Spanish. For instance, for a child
whose degree of exposure is 80\% for Catalan and 20\% for Spanish, the
expected \emph{Exposure} score to the Catalan word \emph{cavall}
{[}\emph{horse}{]}---with a lexical frequency of ---would be , while
that of its translation to Spanish \emph{caballo} would be .

@@

Data processing and visualisation was done in R
(\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
version 4.2.2) using the Tidyverse family of packages
(\protect\hyperlink{ref-wickham2019welcome}{Wickham et al., 2019}). We
implemented the model using \texttt{brms}
(\protect\hyperlink{ref-burkner2017brms}{Bürkner, 2017}), a R interface
to the Stan probabilistic language (2.32.1)
(\protect\hyperlink{ref-carpenter2017stan}{Carpenter et al., 2017}). We
ran four iteration chains with 1,000 iterations each and an additional
1,000 warm-up iterations per chain. Model posterior draws and
predictions were handled using the \texttt{tidybayes}
(\protect\hyperlink{ref-kay2021tidybayes}{Kay, 2021}) and
\texttt{marginaleffects}
(\protect\hyperlink{ref-arel-bundock2022marginaleffects}{Arel-Bundock,
2022}) R packages.

@@

The model posterior showed adequate chain convergence diagnostics for
all parameters (\(\hat{R} \le 1.06\), \(0.022 \le N_{eff}/N < 1.1001\))
and little evidence of multicollinearity. Table~\ref{tbl-coefs} shows
the summary of the posterior distribution of the fixed regression
coefficients, and their degree of overlap with the ROPE. For
interpretability, we report the estimated regression coefficients
transformed to the probability scale\footnote{The logit and probability
  scales relate non-linearly. This means that one logit difference is
  not necessarily translated to a unique value in the probability scale.
  For example, the probability of acquisition of a given word might
  increase in 5\% when age increases from 22 to 23 months, the
  probability of acquisition of the same word might only increase in
  0.2\% when age increases from 31 to 32 months. The linear growth of
  the probability of acquisition differs along the logistic curve, and
  therefore deciding the age point at which to report the estimates of
  the regression coefficients in the probability scale is not trivial.
  Following Gelman et al.
  (\protect\hyperlink{ref-gelman2020regression}{2020}), we report the
  maximum value of such coefficient, which corresponds to the linear
  growth (i.e.~derivative) of the logistic curve at the age at which
  most participants were acquiring a given word. This value can be
  approximated by dividing the coefficient in the logit scale by four:
  \(\hat{\beta_j}/4\), where \(\hat{\beta_j}\) is the estimated mean of
  the posterior distribution of coefficient \(j\).}. The resulting
values correspond to the maximum difference in probability of
acquisition (\emph{Comprehension} or \emph{Comprehension and
Production}) that corresponds to a one standard deviation change in each
predictor.

\hypertarget{tbl-coefs}{}
\begin{table}
\caption{\label{tbl-coefs}Posterior distribution of regression coefficients. }\tabularnewline

\centering
\begin{tabular}{lclr>{}c}
\toprule
 & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Intercepts (at 22 months)}}\\
\hspace{1em}Comprehension and Production & 0.438 & {}[0.379, 0.496] & .088 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d082c12c1b.pdf}\\
\hspace{1em}Comprehension & 0.936 & {}[0.92, 0.949] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d08785c6c0e.pdf}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Slopes (upper bound)}}\\
\hspace{1em}Age (+1 SD, 4.87, months) & 0.405 & {}[0.357, 0.451] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d087e6a2c55.pdf}\\
\hspace{1em}Exposure (+1 SD, 1.81) & 0.233 & {}[0.201, 0.268] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d086eae2508.pdf}\\
\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.014, 0.104] & .037 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d084c7d3767.pdf}\\
\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.086, -0.036] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d0815d12d78.pdf}\\
\hspace{1em}Age $\times$ Exposure & 0.071 & {}[0.039, 0.104] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d082a9a1cf2.pdf}\\
\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0, 0.026] & .985 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d083e54715a.pdf}\\
\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.069, -0.046] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d087de616c.pdf}\\
\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.027, -0.01] & .975 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_4d08238a53bc.pdf}\\
\bottomrule
\end{tabular}
\end{table}

@@

@@

>>>>

source of this study. GGC was supported by the a FPI 2020 research
contract (PRE2019-088165) in a project funded by the Spanish Ministry of
Science, Innovation, and Universities (PGC2018-101831-B-I00). DAV was
supported by the European Union's Horizon 2023 research and innovation
program under Marie Skłodowska--Curie Grant (765556) and a postdoctoral
fellowship from the Foundation for Science and Technology of Portugal
(UIDB/00214/2020). NSG was supported by Academia 2019 award from the
Catalan Institution for Research and Advanced Studies (ICREA). We are
grateful to Chiara Santolin, Ege E. Özer, Alicia Franco-Martínez,
Cristina Rodríguez-Prada, and the rest of the Speech Acquisition and
Perception research group for their helpful feedback. We thank Xavier
Mayoral, Silvia Blanch, and Cristina Cuadrado for their technical
support, and Cristina Dominguez and Katia Pistrin for their efforts in
recruiting infants. We also thank all families and infants who
participated in the experiments.}}

@@

08005\\\texttt{\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}}\And

@@

08005\\\texttt{\href{mailto:avila.varela.daniela@gmail.com}{avila.varela.daniela@gmail.com}}\And

@@

28049\\\texttt{\href{mailto:jignaciocastillejo@gmail.com}{jignaciocastillejo@gmail.com}}\And

@@

08005\\\texttt{\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}}

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Consider the example of the Catalan-Spanish cognate translation
equivalent /\textipa{"gat}-\textipa{"ga.to}/ {[}\emph{cat}{]}, that are
phonologically very similar. When the child is exposed to
/\textipa{"gat}/, they will strongly co-activate /\textipa{"ga.to}/ in
parallel. Therefore, this exposure will count as a learning instance for
both co-activated forms. The case of the non-cognate translation
equivalent /\textipa{"gos-"pe.ro}/ {[}\emph{dog}{]} would be different.
Given the low phonological similarity between both word-forms, an
exposure to \textipa{"gos} will result in a weak activation of
/\textipa{"pe.ro}/ leading to such exposure counting as a learning
instance for /\textipa{"gos}/ (which the child was exposed to), but not
for /\textipa{"pe.ro}/. While /\textipa{"gat}-\textipa{"ga.to}/ will
benefit from parallel activation, /\textipa{"gos}-\textipa{"pe.ro}/ will
not. If the child receives linguistic input from one of the languages
more often than from the other, this effect might affect each form of
the cognate translation equivalent differently. For instance, if the
child receives a larger amount of Catalan input than Spanish input, they
will encounter the Catalan form /\textipa{"gat}/ more frequently than
the Spanish form /\textipa{"ga.to}/. Through parallel activation,
/\textipa{"gat}/ will activate /\textipa{"ga.to}/ more often than vice
versa. Ultimately, /\textipa{"ga.to}/ will benefit more strongly from
its cognate status than /\textipa{"gat}/, as it receives additional
learning instances from its translation equivalent more often than
/\textipa{"gat}/.

@@

\centering
\begin{tabular}{lccccl}
\toprule
Semantic category & List A & List B & List C & List D & Examples\\
\midrule
Household items & 31 & 26 & 30 & 25 & clock, video\\
Food and drink & 29 & 26 & 23 & 27 & sausage, yogurt\\
Animals & 26 & 23 & 19 & 25 & panther, tiger\\
Outside & 14 & 13 & 13 & 15 & farm, stone\\
Body parts & 14 & 12 & 11 & 11 & face, finger\\
\addlinespace
Toys & 11 & 11 & 12 & 13 & piano, racket\\
Clothes & 12 & 12 & 10 & 10 & zipper, sandal\\
Vehicles & 9 & 10 & 11 & 10 & helicopter, tractor\\
Colours & 6 & 6 & 6 & 6 & red, green\\
People & 7 & 4 & 6 & 6 & police, babysitter\\
\addlinespace
Furniture and rooms & 4 & 4 & 4 & 4 & corridor, terrace\\
Time & 2 & 2 & 2 & 2 & day, night\\
Adventures & 1 & 1 & 1 & 1 & witch\\
Parts of things & 1 & 1 & 1 & 1 & wheel\\
\midrule
N & 167 & 151 & 149 & 156 & -\\
\bottomrule
\end{tabular}
\end{table}

@@

We collected data for 1,590 words. We restricted the analyses to
responses to nouns (628 items corresponding to other grammatical classes
were excluded, see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
et al., 2020} for a similar approach). We then excluded items with
missing lexical frequency scores (\emph{n} = 269), items that included
more than one lemma (e.g., \emph{mono/mico} {[}monkey{]}, \emph{n} =
48), multi-word items or phrases (e.g., \emph{barrita de cereales}
{[}cereal bar{]}, \emph{n} = 9). Finally, we removed items without a
translation in the other language (\emph{n} = 32). This resulted in a
final list of 604 items, corresponding to 302 Catalan words and their
302 Spanish translations (302 translation equivalents). After collecting
participants' responses, the final dataset consisted of 138,078
observations, each corresponding to a single response of one participant
to one item. Each translation equivalent received a median of 234
responses (\emph{Min} = 106, \emph{Max} = 872) from participants, both
languages pooled together.

Data processing and visualisation was done in R
(\protect\hyperlink{ref-rcoreteam2013language}{R Core Team, 2013},
version 4.2.2) using the Tidyverse family of packages
(\protect\hyperlink{ref-wickham2019welcome}{Wickham et al., 2019}).

@@

We modelled the probability of participants answering each response
category (\emph{No} \textless{} \emph{Understands} \textless{}
\emph{Understands and Says}) using a Bayesian, multilevel, ordinal
regression model. This model allowed us to estimate both item and
participant word-acquisition trajectories, while estimating the effect
of our variables of interest: \emph{Age} (number of months elapsed
between participants' birth date and questionnaire completion),
\emph{Length} (number of phonemes in the X-SAMPA phonological
transcription of the word-form), \emph{Exposure} (a language
exposure-weighted lexical frequency), and \emph{Cognateness} (defined as
the phonological similarity between translation equivalents). We added
these variables as main effects, together with the two-way and three-way
interactions between \emph{Age}, \emph{Exposure}, and
\emph{Cognateness}. Participant-level and item-level random intercepts
and slopes were included where appropriate, according to the structure
of the data (\protect\hyperlink{ref-barr2013random}{Barr et al., 2013}).
We specified a weakly informative prior around the parameters of the
model. Equation~\ref{eq-model} shows a detailed description of the
model.

\begin{equation}\protect\hypertarget{eq-model}{}{
\begin{aligned}
&\textbf{Response}~(k)~\textbf{to word}~i~ \textbf{by participant}~j \\
\text{Response}_{ij} &\sim \text{Cumulative logit}(\theta_{k_{ij}}) \\ 
&\text{where}~k \in \{\text{No} \rightarrow \text{Understands}, \text{Understands} \rightarrow \text{Understands and Says}\}\\
&\textbf{Distribution parameters} \\
\theta_{k_{ij}} &= (\beta_{0_{k}} + u_{0_{i_{k}}} + w_{0_{j_{k}}}) + (\beta_{1} + u_{1_{i}} + w_{1_{j}}) \cdot \text{Age}_{i} + \\
&(\beta_{2} + u_{2_{i}} + w_{2_{j}}) \cdot \text{Length}_{ij} + 
(\beta_{3} + u_{3_{i}} + w_{3_{j}}) \cdot \text{Exposure}_{ij} + & \\
& (\beta_{4} + u_{4_{i}}) \cdot \text{Cognateness}_{ij} + (\beta_{5} + u_{5_{i}} + w_{3_{j}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij}) + & \\
& (\beta_{6} + u_{6_{i}}) \cdot (\text{Age}_{i} \times \text{Cognateness}_{ij}) + & \\
& (\beta_{7} + u_{7_{i}}) \cdot (\text{Exposure}_{ij} \times \text{Cognateness}_{ij}) & \\
& (\beta_{8} + u_{8_{i}}) \cdot (\text{Age}_{i} \times \text{Exposure}_{ij} \times \text{Cognateness}_{ij}) \\
&\text{where:}\\
& u_{1-8_{i}}~\text{: participant-level adjustments} \\
& w_{1-3_{j}}~\text{: TE-level adjustments} \\
&\textbf{Prior} \\
\beta_{0_{k}} &\sim \mathcal{N}(-0.25, 0.5); ~\beta_{1-5} \sim \mathcal{N}(0, 1) \\
\sigma_{u_{0-8}, w_{0-3}} &\sim \mathcal{N_{+}}(1, 0.25); ~\rho_{u_{0-8}, w_{0-3}} \sim \text{LKJcorr}(2) \\
&\text{where }~\rho_{u_{0-8}, w_{0-3}}~\text{are the correlations between group-level adjustments}\\
\end{aligned}
}\label{eq-model}\end{equation}

@@

We developed the \emph{Exposure} predictor to account for the fact that
bilinguals' exposure to a given word-form is not only a function of the
word-forms lexical frequency, but also of the quantitative input they
receive from the language such word-form belongs to, we expressed
lexical frequencies as the product between both variables. First, we
extracted the child-directed lexical frequency of each word from the
CHILDES database
(\protect\hyperlink{ref-macwhinney2000childes}{MacWhinney, 2000};
\protect\hyperlink{ref-sanchez2019childesdb}{Sanchez et al., 2019}).
Using the corresponding lexical frequencies directly from Catalan and
Spanish was not possible due to the low number of Catalan participants
and tokens available in their corresponding CHILDES corpora, so they
were extracted from the English corpora instead. We then mapped the
lexical frequencies of the English words to their Catalan and Spanish
translations (see \protect\hyperlink{ref-fourtassi2020growth}{Fourtassi
et al., 2020} for a similar approach), and transformed them to Zipf
scores (\protect\hyperlink{ref-vanheuven2014subtlexuk}{Van Heuven et
al., 2014}; \protect\hyperlink{ref-zipf1949human}{Zipf, 1949}). We then
multiplied the resulting lexical frequencies by the reported degree of
exposure of the child to Catalan or Spanish. For instance, for a child
whose degree of exposure is 80\% for Catalan and 20\% for Spanish, the
expected \emph{Exposure} score to the Catalan word \emph{cotxe}
{[}\emph{car}{]}---with a lexical frequency of 6.33---would be 5.06,
while that of its translation to Spanish \emph{coche} would be 1.27.

@@

We implemented the model using \texttt{brms}
(\protect\hyperlink{ref-burkner2017brms}{Bürkner, 2017}), a R interface
to the Stan probabilistic language (2.32.1)
(\protect\hyperlink{ref-carpenter2017stan}{Carpenter et al., 2017}). We
ran four iteration chains using the by-default No U-Turn Sampler
algorithm with 1,000 iterations each and an additional 1,000 warm-up
iterations per chain. Model posterior draws and predictions were handled
using the \texttt{tidybayes}
(\protect\hyperlink{ref-kay2021tidybayes}{Kay, 2021}) and
\texttt{marginaleffects}
(\protect\hyperlink{ref-arel-bundock2022marginaleffects}{Arel-Bundock,
2022}) R packages.

@@

Chain convergence diagnostics showed adequate values, and the model
posterior showed little evidence of correlation, as indicated by
negligible pairwise correlation between the marginal posterior
distribution of coefficients. Table~\ref{tbl-coefs} shows the summary of
the posterior distribution of the fixed regression coefficients, and
their degree of overlap with the ROPE. For interpretability, we report
the estimated regression coefficients transformed to the probability
scale\footnote{The logit and probability scales relate non-linearly.
  This means that one logit difference is not necessarily translated to
  a unique value in the probability scale. For example, the probability
  of acquisition of a given word might increase in 5\% when age
  increases from 22 to 23 months, the probability of acquisition of the
  same word might only increase in 0.2\% when age increases from 31 to
  32 months. The linear growth of the probability of acquisition differs
  along the logistic curve, and therefore deciding the age point at
  which to report the estimates of the regression coefficients in the
  probability scale is not trivial. Following Gelman et al.
  (\protect\hyperlink{ref-gelman2020regression}{2020}), we report the
  maximum value of such coefficient, which corresponds to the linear
  growth (i.e.~derivative) of the logistic curve at the age at which
  most participants were acquiring a given word. This value can be
  approximated by dividing the coefficient in the logit scale by four:
  \(\hat{\beta_j}/4\), where \(\hat{\beta_j}\) is the estimated mean of
  the posterior distribution of coefficient \(j\).}. The resulting
values correspond to the maximum difference in probability of
acquisition (\emph{Comprehension} or \emph{Comprehension and
Production}) that corresponds to a one standard deviation change in each
predictor.

\hypertarget{tbl-coefs}{}
\begin{table}
\caption{\label{tbl-coefs}Posterior distribution of regression coefficients. Median: median of the
posterior distribution in the probability scale. 95\% HDI: 95\% highest
density interval of the distribution. p(ROPE): overlap between the 95\%
HDI and the ROPE, indicating the posterior probability that the true
value of the coefficient is equivalent to zero. In the right column, we
show a graphical representation of the median and 95\% HDI of each
coefficient, with a reference dotted line indicating the location of
zero. }\tabularnewline

\centering
\begin{tabular}{lclr>{}c}
\toprule
 & $\beta$ & 95\% HDI & $p(\text{ROPE})$ & \\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Intercepts (at 22 months)}}\\
\hspace{1em}Comprehension and Production & 0.438 & {}[0.379, 0.496] & .088 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_2718113c70f2.pdf}\\
\hspace{1em}Comprehension & 0.936 & {}[0.92, 0.949] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_2718cf068d3.pdf}\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Slopes (upper bound)}}\\
\hspace{1em}Age (+1 SD, 4.87, months) & 0.405 & {}[0.357, 0.451] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27186be06e3b.pdf}\\
\hspace{1em}Exposure (+1 SD, 1.81) & 0.233 & {}[0.201, 0.268] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27186d8d1719.pdf}\\
\hspace{1em}Cognateness (+1 SD, 0.26) & 0.058 & {}[0.014, 0.104] & .037 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271852a346a4.pdf}\\
\hspace{1em}Length (+1 SD, 1.56 phonemes) & -0.062 & {}[-0.086, -0.036] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27181f2174e.pdf}\\
\hspace{1em}Age $\times$ Exposure & 0.071 & {}[0.039, 0.104] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271847bb40f3.pdf}\\
\hspace{1em}Age $\times$ Cognateness & 0.014 & {}[0, 0.026] & .985 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27184e7d1f57.pdf}\\
\hspace{1em}Exposure $\times$ Cognateness & -0.057 & {}[-0.069, -0.046] & .000 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_271812b03ccf.pdf}\\
\hspace{1em}Age $\times$ Exposure $\times$ Cognateness & -0.018 & {}[-0.027, -0.01] & .975 & \cellcolor[HTML]{f2f2f2}{}\includegraphics[width=1in, height=0.25in]{C:/Users/u155880/Documents/cognate-beginnings/manuscript_files/figure-latex/pointrange_27185801653e.pdf}\\
\bottomrule
\end{tabular}
\end{table}

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

@@

\leavevmode\vadjust pre{\hypertarget{ref-arel-bundock2022marginaleffects}{}}%
Arel-Bundock, V. (2022). \emph{Marginaleffects: Marginal effects,
marginal means, predictions, and contrasts}.
\url{https://CRAN.R-project.org/package=marginaleffects}
